---
title: "Functions"
author: ""
date: ""
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE, error = TRUE,
                      fig.align = "center")
```

## Main ideas

- Create functions

- Function scope and environments

- Function best practices

## Packages

```{r packages}
library(tidyverse)
```

# Notes

## Fundamentals

At a basic level, a function takes inputs and provides an output.
A function is comprised of arguments (formals), body, and environment. The first
two will be our main focus as we use and develop these objects.

```{r example_function}
f <- function(x, y, z) {
  x ^ (y ^ z)
}
```

```{r example_function_call}
f(x = 2, y = 2, z = 3)
```

- Create a function similar to how you create other R objects

- You must use the keyword `function()`

- Wrap the body in braces

- Spacing does not matter, but follow they syntax above for best practices

```{r function_components}
formals(f)
body(f)
environment(f)
```

### Exiting

Most functions end by returning a value (implicitly or explicitly) or in error.

**Implicit return**

```{r implicit_example}

```

**Explicit return**

```{r explicit_example}

```

R functions can return any object.

---

### Calls

Function calls involve the function's name and, at a minimum, values to
its required arguments. Arguments can be given values by

(1) position

```{r call_position}
z <- 1:30
mean(z, .3, FALSE)
```

(2) name

```{r call_name}
mean(x = z, trim = .3, na.rm = FALSE)
```

(3) partial name matching

```{r call_partial}
mean(x = z, na = FALSE, t = .3)
```

The best choice is

```{r call_best}

```

Leave the argument's name out for the commonly used (required) arguments, and 
always specify the argument names for the optional arguments.

## Function scope

R uses lexical scoping. This provides a lot of flexibility, but it can also
be problematic if a user is not careful. Let's see if we can get an idea of
the scoping rules.

```{r scope_example_1, eval=FALSE}
y <- 1

f <- function(x){
  
  y <- x ^ 2
  return(y)
  
}

f(x = 3) 
y 
```

What is the result of `f(x = 3)` and `y` highlighted in yellow?



```{r scope_example_2, eval=FALSE}
y <- 1
z <- 2

f <- function(x){
  
  y <- x ^ 2
  
  g <- function() {
    
    c(x, y, z)
    
  } # closes body of g()
  
  g()
  
} # closes body of f()


f(x = 3) 
c(y, z)  
```

What is the result of `f(x = 3)` and `c(y, z)` highlighted in yellow?

R first searches for a value associated with a name in the current environment.
If the object is not found the search is widened to the next higher scope.

## Function forms

|     Form    |          Description         |         Example(s)        |
|:------------|:-----------------------------|:--------------------------|
| Prefix      | name comes before arguments  | `log(x, base = exp(1))`   |
| Infix       | name between arguments       | `+`, `%>%`, `%/%`         |
| Replacement | replace values by assignment | `names(x) <- c("a", "b")` |
| Special     | all others not defined above | `[[`, `for`, `break`, `(` |

## Help

To get help on any function, type `?fcn_name` in your console, where `fcn_name`
is the function's name. For infix, replacement, and special functions you
will need to surround the function with backticks.

```{r help_examples, eval=FALSE}
?sd
?`for`
?`names<-`
?`%/%`
```

Using function `help()` is an alternative to `?`.

## Function best practices

- Write a function when you have copied code more than twice.

- Try to use a verb for your function's name.

- Keep argument names short but descriptive.

- Add code comments to explain the "why" of your code.

- Link a family of functions with a common prefix: `pnorm()`, `pbinom()`, 
  `ppois()`.
  
- Be consistent in your naming style. A function is an object. If you use
  `snake_case` for all other objects, use `snake_case` for your function
  names.

## Examples

(1) Write a function with an argument `p` that returns the odds. Assume users
    will pass in values for `p` between 0 and 1 inclusive.
    
```{r odds_fcn}

```
    
(2) Write a function called `logit()` with an argument `p` that returns the 
    log odds. Assume users will pass in values for `p` between 0 and 1 
    inclusive.
    
```{r logit_fcn}

```
    
(3) Write a function called `logit_inverse()` with an argument `x` that returns
    the probability value.
    
```{r logit_inverse_fcn}

```

## Practice

(1) Write a function that takes a numeric vector as an input and computes the
    cumulative mean. Function `cumsum()` may be helpful.
    
(2) Write a function that takes a numeric vector as an input and returns the
    mean and standard error. Use an explicit return.
    
(3) Write a function to compute the area of your favorite shape. Have arguments
    necessary to compute the area.