---
title: "Functions"
author: ""
date: ""
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE, error = TRUE,
                      fig.align = "center")
```

## Main ideas

- Create functions

- Function scope and environments

- Function best practices

## Packages

```{r packages}
library(tidyverse)
```

# Notes

## Fundamentals

At a basic level, a function takes inputs and provides an output.
A function is comprised of arguments (formals), body, and environment. The first
two will be our main focus as we use and develop these objects.

```{r example_function}
f <- function(x, y, z) {
  x ^ (y ^ z)
}
```

```{r}
g <- function(dog, cat, fish) {
  dog ^ (cat ^ fish)
}
```

```{r example_function_call}
f(x = 2, y = 2, z = 3)
```

```{r}
g(dog = 2, cat = 2, fish = 3)
```

- Create a function similar to how you create other R objects

- You must use the keyword `function()`

- Wrap the body in braces

- Spacing does not matter, but follow they syntax above for best practices

```{r function_components}
formals(f)
body(f)
environment(f)
```

### Exiting

Most functions end by returning a value (implicitly or explicitly) or in error.

**Implicit return**

```{r implicit_example}
centers <- function(x) {
  c(mean(x), median(x))
}
```

```{r}
centers(1:10)
centers(c(pi, 0, -5, 100, log(40)))
```

```{r}
centers <- function(x) {
  x_bar <- mean(x)
  med <- median(x)
  #c(mean(x), median(x))
  result <- c(x_bar, med)
}
```

**Explicit return**

```{r explicit_example}
standardize <- function(y) {
  result <- (y - mean(y)) / sd(y)
  return(result)
}
```

```{r}
standardize(y = 1:10)
standardize(y = c(2, 4, 6, 1, -8, pi))
```

```{r}
mtcars %>% 
  mutate(mpg_stand = standardize(mpg))
```

R functions can return any object.

### Calls

Function calls involve the function's name and, at a minimum, values to
its required arguments. Arguments can be given values by

(1) position

```{r call_position}
z <- 1:30
mean(z, .3, FALSE)
```

(2) name

```{r call_name}
mean(x = z, trim = .3, na.rm = FALSE)
```

(3) partial name matching

```{r call_partial}
mean(x = z, na = FALSE, t = .3)
```

The best choice is

```{r call_best}
mean(z, trim = 0.3, na.rm = TRUE)
```

Leave the argument's name out for the commonly used (required) arguments, and 
always specify the argument names for the optional arguments.

## Function scope

R uses lexical scoping. This provides a lot of flexibility, but it can also
be problematic if a user is not careful. Let's see if we can get an idea of
the scoping rules.

```{r scope_example_1, eval=FALSE}
y <- 1

f <- function(x){
  b <- 5 # we can't access this in our global environment
  y <- x ^ 2
  return(y)
  
}

f(x = 3) 
y 
```

What is the result of `f(x = 3)` and `y` highlighted in yellow?


```{r scope_example_2, eval=FALSE}
y <- 1
z <- 2

f <- function(x){
  
  y <- x ^ 2
  
  g <- function() {
    
    c(x, y, z)
    
  } # closes body of g()
  
  g()
  
} # closes body of f()


f(x = 3) 
c(y, z)  
```

What is the result of `f(x = 3)` and `c(y, z)` highlighted in yellow?

R first searches for a value associated with a name in the current environment.
If the object is not found the search is widened to the next higher scope.

## Function forms

|     Form    |          Description         |         Example(s)        |
|:------------|:-----------------------------|:--------------------------|
| Prefix      | name comes before arguments  | `log(x, base = exp(1))`   |
| Infix       | name between arguments       | `+`, `%>%`, `%/%`         |
| Replacement | replace values by assignment | `names(x) <- c("a", "b")` |
| Special     | all others not defined above | `[[`, `for`, `break`, `(` |

## Help

To get help on any function, type `?fcn_name` in your console, where `fcn_name`
is the function's name. For infix, replacement, and special functions you
will need to surround the function with backticks.

```{r help_examples, eval=FALSE}
?sd
?`for`
?`names<-`
?`%/%`
```

Using function `help()` is an alternative to `?`.

## Function best practices

- Write a function when you have copied code more than twice.

- Try to use a verb for your function's name.

- Keep argument names short but descriptive.

- Add code comments to explain the "why" of your code.

- Link a family of functions with a common prefix: `pnorm()`, `pbinom()`, 
  `ppois()`.
  
- Be consistent in your naming style. A function is an object. If you use
  `snake_case` for all other objects, use `snake_case` for your function
  names.

## Examples

(1) Write a function with an argument `p` that returns the odds. Assume users
    will pass in values for `p` between 0 and 1 inclusive.
    
```{r odds_fcn}
compute_odds <- function(p) {
  stopifnot(p >= 0, p < 1)
  
  p / (1 - p)
}
```

```{r}
compute_odds(.9)
compute_odds(.4)
compute_odds(c(0, .5, .75, .99))
compute_odds(1)
compute_odds(-1)
```
    
(2) Write a function called `logit()` with an argument `p` that returns the 
    log odds. Assume users will pass in values for `p` between 0 and 1 
    inclusive.
    
```{r logit_fcn}
logit <- function(p) {
  log(compute_odds(p))
}
```

```{r}
logit(0.5)
logit(-4)
logit(10)
logit(.10)
```
    
(3) Write a function called `logit_inverse()` with an argument `x` that returns
    the probability value.
    
```{r logit_inverse_fcn}
logit_inverse <- function(x) {
  stopifnot(is.numeric(x), is.finite(x))
  
  exp(x) / (1 + exp(x))
}
```

```{r}
logit_inverse("a")
logit_inverse(Inf)
```

```{r}
logit_inverse(10)
logit_inverse(-60)
```

## Practice

(1) Write a function that takes a numeric vector as an input and computes the
    cumulative mean. Function `cumsum()` may be helpful.
    
```{r}
cummean <- function(x) {
  cumsum(x) / (1:length(x))
}
```

```{r}
cummean(c(-2, -1, 0, 1, 2))
```

    
(2) Write a function that takes a numeric vector as an input and returns the
    mean and standard error. Use an explicit return.
    
```{r}
y <- c(8, 0, -19, NA, NA, 3, 4)
length(y)

y[!is.na(y)]
```

    
```{r}
summary_stats <- function(x) {
  x_bar <- mean(x, na.rm = TRUE)
  se <- sd(x, na.rm = TRUE) / sqrt(length(x[!is.na(x)]))
  
  result <- c(x_bar, se)
  return(result)
}
```

```{r}
summary_stats(rnorm(10000))
```

(3) Write a function to compute the area of your favorite shape. Have arguments
    necessary to compute the area.
    
```{r}
trapezoid <- function(base_1, base_2, height) {
  ((base_1 + base_2) / 2) * height
}
```

